---
title: Finches data analysis example
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the packages

Loading the R packages.

```{r}

library(readxl)     
library(tidyverse) 
```

Read the finches data
```{r}
finches <- read_excel("finches_data.xlsx")

```

Histogram of beak length, grouped bt survival, with labels
```{r}
ggplot(
  data = finches,                    
  mapping = aes(x = beak_length,      
                fill = outcome)       
) +
  geom_histogram(bins = 14) +          
  facet_wrap(~ outcome, ncol = 1) +    
  guides(fill = FALSE) +               
  labs(
    title = "Figure 1.",              
    x = "Beak Length (mm)",           
    y = "Number of Birds"             
  )

```

Summarize the data by outcome (survived vs. died)
```{r}
beak_length_grouped_summary <- 
  finches %>% 
  group_by(outcome) %>% 
  summarize(mean = mean(beak_length),
            sd = sd(beak_length),
            n = n()) %>% 
  mutate(sem = sd / sqrt(n),
         upper = mean + 1.96 * sem,
         lower = mean - 1.96 * sem)

```

Bar chart of mean beak lengths
```{r}
ggplot(
  data = beak_length_grouped_summary,   
  mapping = aes(x = outcome,            
                y = mean,               
                fill = outcome)         
) +
  geom_col() +                          
  geom_errorbar(                      
    mapping = aes(ymin = lower,         
                  ymax = upper),        
    width = .3                          
  ) +
  guides(fill = FALSE) +                
  labs(
    title = "Figure 2.",                
    x = "Survival Outcome",             
    y = "Beak Length (mm)"              
  )
```

Get a vector of beak lengths for birds that died
```{r}
beak_length_died <-
  finches %>%                     
  filter(outcome == "died") %>%   
  pull(beak_length)               
```

Print the new object in the console... it is a vector
```{r}
beak_length_died
 
```

Get a vector of beak lengths for birds that survived
```{r}
beak_length_survived <-
  finches %>% 
  filter(outcome == "survived") %>% 
  pull(beak_length)
```

Print the results in the console 
```{r}
beak_length_survived

```

Perform a two-sample t-test assuming unequal variances
```{r}
t.test(beak_length_died, beak_length_survived)

```



